Empirical Bayes analysis of \RNAseq{} data for detection of gene expression heterosis

### 2015-08-18 response.tex

\begin{document}

Editor comments:

\comment{The proposed model appears to be an extension of Ji et al. (2014) albeit with a differently specified likelihood function (negative binomial or Poisson-gamma) as perceived to be more appropriate for count data relative to the Gaussian likelihood function pursued in Ji et al. (2014). Both reviewers indicate that they are little underwhelmed by the performance of your proposed procedure relative to some of the existing methods that you compare against (i.e., ShrinkBayes, edgeR and baySeq).  Furthermore, they seem somewhat concerned about an ‘apples vs. oranges’ types of comparisons since the proposed hypothesis testing strategy in the proposed method is really not an option in the other methods.}

\todo{We don't have artificially large signals. I didn't perceive the ``apples vs. oranges'' comparison even though I agree.}

@@

\comment{c.      Right before Section 2.2…it is assumed apriori independence, correct?}

\response{Yes, added \emph{a priori}.}

\comment{d.      Page 4: You have a two-step approach where you estimate the parameters of interest to get estimates of the hyperparameters and then restimate.  How much change is there is these estimates?  Why aren't the methods of McCarthy used as a competitor?}

\response{We have added a figure (Fig 3) to compare gene-specific parameter estimates and a paragraphing describing the comparison in Section 4. We aren't exactly sure which McCarthy methods you are referring to. We already use an approach based on edgeR and the point estimates themselves do not naturally lead to a statistic for determining the plausibility of heterosis.}

@@

\comment{e.      Page 11 "These figures show marked departures from ……"}

\response{Fixed.}

\comment{f.      Page 11 : Does it really matter that the posterior is different from the prior?  One often assume aprior independence but expects there to be posterior correlation.  Perhaps I'm just missing something in this discussion.}

\todo{Due to the nature of an empirical Bayes analysis, we are learning nothing about the distribution of the gene-specific parameter estimates other than through the hyperparameters. Yes, we can learn something about the parameters for a particular gene, but the information content in the data is very poor, i.e. only 12 total observations to infer about 4 parameters. Had we performed a fully Bayesian analysis, we could have learned more about the distribution of the gene-specific parameter estimates via the joint distributions of the hyperparameters (we believe this is what you are referring to). We believe that improved estimation of the gene-specific parameter distribution should lead to improved ability to detect heterosis.}

### 2015-08-24 ebayes_rnaseq_heterosis.tex
updated fig caption

For the methods in Sections \ref{s:ebayes}, \ref{s:shrinkbayes}, and \ref{s:alternative}, we sorted genes according to the computed measure of the strength of evidence for heterosis. From these sorted lists, we constructed receiver-operating characteristic (ROC) curves to evaluate the ability of these methods to distinguish genes with heterosis, as defined in Section \ref{s:sim_data}, 
% From Dan, "why this citation?"
from those without heterosis \citep{landau2013dispersion}. A representative set of ROC curves is shown in Figure \ref{f:roc}. 
\begin{figure}[htbp]
\centerline{\includegraphics[width=\textwidth]{exampleROC0_1}}
\caption{Representative ROC curves for false positive rates below 0.1 for the approaches using \edgeR{} and \baySeq{} described in Section \ref{s:alternative}, the \ShrinkBayes{} approach described in Section \ref{s:shrinkbayes} and our eBayes approach described in Section \ref{s:ebayes}.}
\label{f:roc}
\end{figure}
For this simulation, we can see that the approaches based most closely on the model in Section \ref{s:model}, i.e. eBayes and \ShrinkBayes{}, provide the best true positive rate for a given false positive rate. Also, as expected, as the sample size increases, our ability to distinguish genes with heterosis from genes without improves.

>>>>

For the methods in Sections \ref{s:ebayes}, \ref{s:shrinkbayes}, and \ref{s:alternative}, we sorted genes according to the computed measure of the strength of evidence for heterosis. From these sorted lists, we constructed receiver-operating characteristic (ROC) curves to evaluate the ability of these methods to distinguish genes with heterosis, as defined in Section \ref{s:sim_data}, 
% From Dan, "why this citation?"
from those without heterosis \citep{landau2013dispersion}. A representative set of ROC curves is shown in Figure \ref{f:roc}. 
\begin{figure}[htbp]
\centerline{\includegraphics[width=\textwidth]{exampleROC0_1}}
\caption{Representative ROC curves for false positive rates below 0.1 for the approaches using \edgeR{} and \baySeq{} described in Section \ref{s:alternative}, the approach of \cite{} using log-counts, the \ShrinkBayes{} approach described in Section \ref{s:shrinkbayes} and the eBayes approach described in Section \ref{s:ebayes} using both normal and laplace distributions for the half-parental difference and hybrid effect.}
\label{f:roc}
\end{figure}
For this simulation, we can see that the approaches based most closely on the model in Section \ref{s:model}, i.e. eBayes and \ShrinkBayes{}, provide the best true positive rate for a given false positive rate. Also, as expected, as the sample size increases, our ability to distinguish genes with heterosis from genes without improves.
